---
title: "Figure 2A"
author: "Vlada Milchevskaya"
#date: "11.11.20"
output: 
    prettydoc::html_pretty:
        toc: true
        theme: cayman
        highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("animalBreeding")
library(ggplot2)
library(ggpubr)
library(fields)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
source("../helper_function_fig2.R")
```


# Fig. 2A

We calculate the minimal required number of breeding (Y-axis) needed to obtain 
the desired number of offsprings (X-axis) with 90% confidence, using three methods: 

1. The naïve ratio (see Suppl.), 

2. the gold standard textbook model suggested by Festing, 

3.  and our method denoted as Poisson.

We use the following values for the fertility, average litter size and the required confidence:
```{r, eval = TRUE, include=TRUE}
effective_fertility <- 0.7
average_litter_size <-  7
required_confidence <-  0.9
```
We further calculate the difference in required breedings between 
the textbook method (Festing) and our method, expressed in percentage 
relative to the gold-standard method (Festing): 
This is the number of "saved" breedings.


```{r, eval = FALSE}
# data generation, old names kept till the end, then changed
req_pups <- seq(2,100,1)
# see ../helper_function_fig2.R for the function
data4b_prelim <- calc_data4fig_2A(
    req_pups,
    genotype_probability = 1, 
    confiP = required_confidence, 
    effertyP = effective_fertility, 
    litmean = average_litter_size)

# % relative to Festing
animals_saved_fp_percent <- round(
    100*(data4b_prelim$req_festing-data4b_prelim$req_poisson)/data4b_prelim$req_festing)

# % relative to Poisson
animals_saved_pf_percent <- round(
    100*(data4b_prelim$req_festing-data4b_prelim$req_poisson)/data4b_prelim$req_poisson)

data4b <- data.frame(
    data4b_prelim, 
    animals_saved_fp_percent, 
    animals_saved_pf_percent)
data2a <- data4b
save(data2a, file = "data2a.Rdata")
```



```{r}
load("data2a.Rdata")


plot2A_hist <- ggplot(data2a,
                      aes(x=req_pups, y=animals_saved_pf_percent))+
  geom_bar(stat="identity",col="darkgrey")  + 
  theme_pubr() + 
  #theme(axis.title.x = element_blank(), axis.text.x = element_blank()) + 
  ylab("Percent saved \n animals")

ggsave(plot2A_hist, 
       filename = "plot2A_hist_annotated.pdf", width = 7, height = 5)

plot2A_points <- 
  ggplot(data = data2a , aes(x=req_pups, y = animals_saved_fp_percent)) + 
  #geom_point(aes(y=req_br_festing50, col = "Festing 50%"), colour = "red", alpha=0.5, size=0.5) + 
  #geom_line(aes(y=req_br_festing50, col = "Festing 50%"), alpha=0.3, size=0.3) + 
  #geom_point(aes(y=req_br_festing90, col = "Festing 90%")) +
  geom_point(aes(y= req_poisson, col = "poisson")) + 
  geom_point(aes(y=req_festing, col = "festing")) + 
  geom_point(aes(y=req_mendel, col = "mendel")) + 
  labs(x="Required number of pups",
       y = "Smallest number \n of breedings",
       caption = "Mean litter size: 7,  Effective fertility: 70%",
       title = "Model comparison for the smallest required number of breegins", colour = "") +
  #labs(x='',y='',title='',colour = "Method") +
  theme_pubr() +  theme(legend.position = c(0.85, 0.25))

ggsave(plot2A_points, filename = "plot2A_points_annotated.pdf", width = 7, height = 7)


plot2A <- ggarrange(
  plotlist = list(plot2A_hist, plot2A_points), ncol = 1, 
  heights = c(1,3))
plot2A
ggsave(plot2A, filename = "plot2A_annotated.pdf", width = 7, height = 7)
save(plot2A_hist, plot2A_points, file = "gg_plots_2A.Rdata")
```

