---
title: "Package 'animalBreeding'"
output:
  rmarkdown::html_vignette:
    toc: true
#output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
library(ggplot2)
```

# Setup

For installation, run the code below.

```{r libraries, eval = FALSE}
library(devtools)
```

```{r setup, message=FALSE, comment=FALSE, cache=FALSE, results=FALSE}
library(animalBreeding)
```

At the moment, the package repository is private, therefore
the installation will only work with the `auth_token` argument. 
Later it will be replaced.

# Usage (simple)

## Offsprings of a single genotype

When the breedins set up is aimed at a offsprings of single genotype, 
or simply a total number of born pups, one may use the following function to 
calculate the required number of breedings. 

Here we calculate how many breedings are required to ensure that upon mating FVB/N mice, 
there are at least 20 pups, born within a period of 3 days, with the success probability of 90%.
To run the calculation, we also assume that the average number of pups in a single litter is 6.



```{r}
n_breedings <- singleGenotype(
  confidence_p = 0.95,
  birth_days = 3,
  n_offsprings = 20,
  strain = "FVB/N",
  litter_mean = 6)

print(n_breedings)
```

## Offsprings of multiple genotypes

Assume we breed +/- and +/- mice. According to Mendel law, the probabilities of 
offsprings +/+, +/- and -/- are 25%, 50% and 25%. 
We only need 10 pups of -/- genotype, and require them to be born within a perios of 3 days. 
Further, we want all pups to be of one and the same sex (either male or female).

Lastly, we want to ensure that the above requirements are met in at least 95 out of 100 trials.


```{R}
n_breedings <- multiGenotype(
  confidence_p = 0.95, 
  birth_days = 3, 
  genotypes_N = c(0,0,10), 
  genotypes_p = c(0.25, 0.5, 0.25),
  sex_distribution = "unimportant", 
  strain = "FVB/N",
  litter_mean = 6)

n_breedings
```

One may want to know how many animals (in total) should be expected to be born with the breeding setup from above:
```{r}
expectBorn(n_breedings)
```




# Usage (not as simple, will be deprecated?)

For instance, we need 200 offsprings, and we know the following: 

- The average litter size of a fertile mouse is 7

- The effective fertility of a mouse is 70%.

- The desider number of offsprings is 200, and the desired confidence is 95%

Then, the following arguments should be provided:

```{r festing as example}
calculate_needed_breedings(
    confidence_p = 0.95, 
    effective_fertility_p = 0.65, 
    n_needed = 200, 
    litter_mean = 7, 
    litter_sd = 2.5, # default, in accordance with the Festing book
    method = "festing" # or "poisson", "binomial", "empirical"
    )
```
## Festing model in details

### Reproducing Table 3.12

To reproduce Table 3.12 from the "Animal production and breeding" textbook, 
use the `calculate_needed_litters_textbook()` function. For instance, the number of 
litters required to give 97.5% probability of achieving 25 animals, with the 
average litter size being 4, is 11 (or 10 in the textbook).

```{r festing textbook 1}
calculate_needed_litters_textbook(
    confidence_p = 0.975, 
    litter_mean = 4, 
    litter_sd = 2.5,
    n_needed = 25)
```

Note that the calculation above uses values for the quantiles of the Gaussian 
distribution. Quite likely, the authors of the textbook used a different 
approximation from the one I use now, 
which is why our value may be a little different from the one in the textbook. 

In particular: The smallest integer that 
guarantees $P\geq 0.975$ would be 11 in the case above.

### Reproducing Table 3.13

To reproduce Table 3.13 from the "Animal production and breeding" textbook, use 
the `calculate_needed_breedings_textbook()` function. For instance, the number of 
breedings required to give 97.5% probability of achieving more than 15 
litters, with the effective fertility being 20%, is 119 (according to 
the textbook). 

```{r festing textbook 2}
calculate_needed_breedings_textbook(
    confidence_p = 0.975, 
    effective_fertility_p = 0.2, 
    n_litters = 15, 
    textbook_error = TRUE) # default FALSE, we should rename this argument or leave it out..

```

The values I get when implementing the exact same model are slightly 
different, possibly due to the approximations the authors used, 
and the subtle difference between $P(X \leq x)$ and $P(X < x)$.

Thus, for the same figures as above, one would need 113 breedings but not 119.

```{r festing textbook 3}
calculate_needed_breedings_textbook(
    confidence_p = 0.975, 
    effective_fertility_p = 0.2, 
    n_litters = 15)
```

Below I check that the calculation was correct. To confirm that `pbinom` 
calculates $P(X \leq x)$ see the `lower.tail` argument in the `?pbinom` 
description.
```{r festing check}
# probability to get 10 or more litters from 110 breedings
1 - pbinom(q = 14, size = 113, prob = 0.2)

# probability to get 10 or more litters from 109 breedings
1 - pbinom(q = 14, size = 112, prob = 0.2)

# thus, 110 is the minimal required number of breedings to have P>0.975
```



## Various methods (examples)

To calculate the number of breedings directly from the required number of 
offsprings, run the following:

```{r various methods}
# should be close to poisson, we may remove this method later on
calculate_needed_breedings(
    confidence_p = 0.95, 
    effective_fertility_p = 0.6, 
    n_needed = 200, 
    litter_mean = 7, 
    method = "binomial"
    )

# the main alternative
calculate_needed_breedings(
    confidence_p = 0.95, 
    effective_fertility_p = 0.6, 
    n_needed = 200, 
    litter_mean = 7, 
    method = "poisson"
    )


# should be the same as poisson, because I provided an "empirical" observation
# generated from a Poisson distribution
calculate_needed_breedings(
    confidence_p = 0.95, 
    effective_fertility_p = 0.6, 
    n_needed = 200, 
    offsprings_n_sample = rpois(n = 1000, lambda = 7),
    method = "empirical"
    )

# textbook
calculate_needed_breedings(
    confidence_p = 0.95, 
    effective_fertility_p = 0.6, 
    n_needed = 200, 
    litter_mean = 7, 
    litter_sd = 2.5,
    method = "festing"
    )

```


## Genotype-specific calculations

Assume you need no less than 10 KO, 15 heterozygous and 5 WT offsprings to be 
born, with the confidence of 95%. Effective fertility of the mice is 70%, and 
the average litter size is 4. The parameter `genotypes_p=c(0.25, 0.5, 0.25)` 
reflects standard Mendelian probablities for the respective genotypes, 
and `genotypes_N = c(1,0,0)` specifies the desired number of offsprings.

This is only implemented for the Poisson model at the moment.
For the Festing model,unfortunately, this is not possible (since Normal 
distribution cannot be combined with the binomial for the genotype drawing).

```{r several genotypes}
x <- breed_genotype(
    confidence_p = 0.95, 
    effective_fertility_p = 0.7, 
    genotypes_p = c(0.25, 0.5, 0.25), 
    genotypes_N = c(10,15,5),
    litter_mean = 4, 
    method = "poisson")
x


```

<!--
# Methods

First, let us assume that every fertile animal $i$ produces $\xi_i$ offsprings.

Also, let us assume that we conducted $K$ breeding events, out of which $\eta$ were 
successful. According to Festing, $\eta$ follows a Binomial distribution defined 
by the effective fertility of the animals. Namely, if the effective fertility 
is $65\%$, then $p_{fert}=0.65$ and
$$\eta \sim Binom(p_{fert}, K).$$

Then, the total number of offsprings $\chi$ would be
$$ \sum\limits_{i=1}^{\eta} \xi_i,$$
where $\xi_i$ are i.i.d random variables.

Unfortunately, we do not know the distribution of $\xi_i$, but we could make 
some assumptions about it from some apriori knowledge. 

Festing, for instance, assumes that $\xi_i$ is Normally distributed:
$$\xi_i \sim N(\mu, \sigma=2.5),$$
where $\mu$ is the average number of offsprings an animal produces, and the 
variance $\sigma$ is fixed to $2.5$. 

Alternatively, one could assume $\xi_i$ to come from Poisson or Binomial 
distributions, or even estimate it empirically from the some 
previously produced data. 

--> 

