---
title: "Package 'animalBreeding'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

For installation from GitHub, run the code below. 
Since the package repository is private at the moment, 
you need the `auth_token` argument.

```{r setup, message=FALSE}
library(devtools)
install_github("IMSBComputationalBiology/animalBreedingProj", 
               subdir="animalBreeding",
               auth_token = "f6024685998b779a9db258994a7b4e44468b754c", 
               force=TRUE)
library(animalBreeding)
```

# Usage

To reproduce the Table 3.12 from the "Animal production and breeding" textbook, 
use the `calculate_needed_litters()` function. For instance, the number of 
litters required to give 97.5/% probability of achieving 25 animals, with the 
average litter size being 4, is 10.

```{r}
calculate_needed_litters(condifence_p = 0.975, 
                         litter_mean = 4, 
                         litter_sd = 2.5, 
                         n_needed = 25)
```